### 2017-06-16 paper/images/flowchart/flowchart_beamer_hd.tex
[major] add updated presentation for conference

\documentclass[11pt]{beamer}
\usefonttheme{serif}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric,backgrounds,positioning-plus,node-families,calc}
\tikzset{
  basic box/.style = {
    shape = rectangle,
    align = center,
    draw  = #1,
    fill  = #1!25,
    font  = \Large,
    rounded corners},
  header node/.style = {
    %Minimum Width = header nodes,
    font          = \huge\bf,
    text depth    = +0pt,
    fill          = white,
    draw},
  header/.style = {%
    inner xsep = +1.2em,
    inner ysep = +1.5em,
    append after command = {
      \pgfextra{\let\TikZlastnode\tikzlastnode}
      node [header node] (header-\TikZlastnode) at (\TikZlastnode.north) {#1}
      node [span = (\TikZlastnode)(header-\TikZlastnode)]
        at (fit bounding box) (h-\TikZlastnode) {}
    }
  },
  hv/.style = {to path = {-|(\tikztotarget)\tikztonodes}},
  vh/.style = {to path = {|-(\tikztotarget)\tikztonodes}},
  fat blue line/.style = {ultra thick, blue}
}
\setbeamertemplate{navigation symbols}{}
\setbeamersize{text margin left=0.2em,text margin right=0.2em}

\def \scaling {0.15}
\begin{document}
\begin{frame}
%\hskip -2em
\begin{tikzpicture}[scale=0.52, every node/.style={transform shape}, thick, >=latex]
	\node[basic box = blue, header = Teaching Phase] (teaching) {
		\includegraphics[scale=\scaling]{teaching_hd}\\
		A demonstration is performed by moving the \\
		Baxter arms in the appropriate trajectory
	};

	\node[north right = of teaching, basic box = orange, header = Trajectory Parameterization, shift = (right:1.2*x_node_dist)] (dmp) {
		Baxter Left Arm Trajectory\\
		\includegraphics[scale=\scaling]{dmp_hd}\\
		Recorded trajectory is parameterized by DMP
	};

	\node[east below = of teaching, basic box = green, header = Testing Phase] (testing) {
		Modified Posture\\
		\includegraphics[scale=\scaling]{testing_hd}\\
		DMP can accomodate any posture by \\
		changing goal parameter
	};

	\node[north right = of testing, basic box = red, header = DMP Generalization, shift = (right:1.2*x_node_dist)] (generalization){
		DMP Trajectory\\
		\includegraphics[scale=\scaling]{dmp_new_hd}
	};

    \path[ultra thick, blue, ->](teaching) edge[->] (dmp);
	\path[ultra thick, orange, ->](dmp) edge[->] (testing);
	\path[ultra thick, green, ->](testing) edge[->] (generalization);
\end{tikzpicture}
\end{frame}
\end{document}